\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Checkpoint eval overview

% \TODO{Drop the automated from the title maybe?}
% \nv{Updates}

@@

@@

\begin{figure}[t]
\centering
% https://docs.google.com/drawings/d/13Zx3a9bWi1RvpOgNCSiXirp9UOte-XuaJQBi_lmmPMc/edit
\includegraphics[width=0.4\textwidth]{\detokenize{./figs/pash_lattice.pdf}}
\caption{
  \textbf{Runtime setup lattice.}
  Parallel No Eager and Blocking Eager improve over sequential, but are not directly comparable.
  Eager adds \sys's optimized \ttiny{eager} relay, and \sys uses all primitives in \S\ref{optimizer} (Fig.~\ref{fig:microbenchmarks}).
% Describe. \TODO{Combine this figure with the above code maybe.}
}
\vspace{-15pt}
\label{fig:lattice}
\end{figure}

@@

This section reports on whether \sys can indeed scale shell scripts out automatically and correctly, using several scripts collected out from the wild along with a few micro-benchmarks for targeted comparisons.

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) \todo{49/54} scripts, whereas for the rest 5 the parallel execution time is close to the sequential (0.89, 0.91, 0.94, 099, 1.01$\times$).
In this paragraph we summarize results for \ttt{--width=16}, but the rest of our evaluation section reports on varying \ttt{width}s.
For a subset of our benchmarks~\sx{ours}, we decouple the performance benefits stemming from \sys's DFG transformations alone from the ones stemming from its runtime primitives:
  the compiler's average speedup is \todo{$9.3\times$}, jumping to \todo{$13.2\times$} when runtime optimizations are used.
A different subset ($30$ benchmarks) see an average speedup of $6.02\times$~\sx{unix50}.
% , depending on the classed of commands involved
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see speedups of $2.44\times$ and $12.7\times$.
\sys outperforms \ttt{sort --parallel} \todo{by a significant margin}\footnote{
\nv{Nikos add point about experts}
}, and performs better than GNU \ttt{parallel}, which returns incorrect results if used without care~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

@@

\heading{Parallelism}
The level of parallelism is configured using \sys's \ttt{--width=N}, a flag that determines the number of times that \sys applies its transformations and the resulting width of the data-parallel DFG.
% The value  of \ttt{N} is shown as the x-axis of Fig.\ref{fig:microbenchmarks}. 
We cannot control the initial parallelism of a script (\eg a command could spawn 10 processes), and thus parallel scripts produced by \sys often reach maximum parallelization benefits with a value of \ttt{N} that is smaller than the physical cores of the hardware setup (in our case 64).

@@

We evaluate \sys on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.
The goal is to evaluate
  (i) the performance benefits stemming from \sys's  DFG transformations alone, including how \ttt{--width} affects speedup, 
  (ii) performance benefits stemming from \sys's runtime primitives, showing results for several configurations of the runtime lattice (Fig.~\ref{fig:lattice}).

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for \ttt{width=} 2--64.}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
  For some pairs of configurations, \sys produces identical parallel scripts and thus only one is shown.
}
%   Two types of pairs where x y z coincide
%   - \sys adds no split, whenever whenever the DFG width is constant for the entire pipeline. In these cases the blue line is identical to the red line, which not shown.
%    Conversely, whenever both blue and red lines are shown, split nodes are needed due to changes in the pipeline width and result in performance improvements.
% 
%   - without split, the script is sequential (spell, bigrams)
% Early on they have a pure or no parallelizable pure
% and thus introduce an aggregator resulting in width of 1 right after, and requires a split.
% Thus, any configuration that does not has \ttt{split} is practically sequential beyond that point---\ie No Eager, Blocking Eager, and \sys w/o Split see no speedup.
% 
% Blocking eager: data parallelism, but no task parallelism
% No eager: blocking FIFO does not allow pushing -> underutilize CPU
% 
% The reason why Blocking Eager is usually slower than No Eager is due to \ttt{agg}regators that read multiple input streams at the same time---such \ttt{sort -m}.
% 
% Sort-sort possibly shows the full spectrum:
%   (iii) \sys's \ttt{eager} offers benefits despite the lack of \ttt{split} because it fully parallelizes the first \ttt{sort} while maintaining task parallelism
%   (iv) \sys gets full benefits because \ttt{split}ting allows parallelizing the second \ttt{sort} too.

      % (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      % Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

>>>>

@@

\nv{Make sure we use \ttt{width} everywhere}
\nv{Decide on parallelizable regions}

@@

\begin{figure}[t]
\centering
% https://docs.google.com/drawings/d/13Zx3a9bWi1RvpOgNCSiXirp9UOte-XuaJQBi_lmmPMc/edit
\includegraphics[width=0.4\textwidth]{\detokenize{./figs/pash_lattice.pdf}}
\caption{
  \textbf{Runtime setup lattice.}
  Parallel No Eager and Blocking Eager improve over sequential, but are not directly comparable.
  Eager adds \sys's optimized \ttiny{eager} relay, and \sys uses all primitives in \S\ref{optimizer} (Fig.~\ref{fig:microbenchmarks}).
}
\vspace{-15pt}
\label{fig:lattice}
\end{figure}

@@

This section reports on whether \sys can indeed offer performance benefits automatically and correctly using several scripts collected out from the wild along with a few micro-benchmarks for targeted comparisons.

@@

\heading{Highlights}
In this paragraph we summarize results for \ttt{--width=16}, but \sys's evaluation reports on varying \ttt{width}s.
Overall, \sys accelerates 49/54 scripts by \todo{1.92--61.1}$\times$; % Eurosys: results for 16!!
  for the rest, the parallel performance is comparable to the sequential (0.89, 0.91, 0.94, 099, 1.01$\times$).
For a subset of our benchmarks~\sx{ours}, we decouple the speedup due \sys's DFG transformations alone from the one due to its runtime primitives:
  the former averages \todo{$9.3\times$}, jumping to \todo{$13.2\times$} due to the latter.
A different subset ($30$ benchmarks) sees average speedups of $6.02\times$~\sx{unix50}.
% , depending on the classed of commands involved
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see speedups of $2.44\times$ and $12.7\times$.
\sys significantly outperforms \ttt{sort --parallel}, which involves hand-tuning by its command developers, and performs better than GNU \ttt{parallel}, which returns incorrect results if used without care~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

@@

\heading{Parallelism}
The level of parallelism is configured using \sys's \ttt{--width=W}, a flag that determines the number of times that \sys applies its transformations and the resulting width of the data-parallel DFG.
% The value  of \ttt{N} is shown as the x-axis of Fig.\ref{fig:microbenchmarks}. 
\sys does not control a script's initial parallelism (\eg a command could spawn 10 processes), and thus the resulting scripts often reach maximum parallelization benefits with a value of \ttt{W} smaller than the physical cores of our hardware setup (in our case 64).

@@

We evaluate \sys on popular, common, or classic \unix pipeline patterns~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked}.
The goal is to evaluate performance benefits due to \sys's  
  (i) DFG transformations alone, including how \ttt{--width} affects speedup, 
  (ii) runtime primitives, showing results for several configurations of the runtime lattice (Fig.~\ref{fig:lattice}).

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for \ttt{width=} 2--64.}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
  For some pairs of configurations, \sys produces identical parallel scripts and thus only one is shown.
}
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

%   Two types of pairs where x y z coincide
%   - \sys adds no split, whenever whenever the DFG width is constant for the entire pipeline. In these cases the blue line is identical to the red line, which not shown.
%    Conversely, whenever both blue and red lines are shown, split nodes are needed due to changes in the pipeline width and result in performance improvements.
% 
%   - without split, the script is sequential (spell, bigrams)
% Early on they have a pure or no parallelizable pure
% and thus introduce an aggregator resulting in width of 1 right after, and requires a split.
% Thus, any configuration that does not has \ttt{split} is practically sequential beyond that point---\ie No Eager, Blocking Eager, and \sys w/o Split see no speedup.
% 
% Blocking eager: data parallelism, but no task parallelism
% No eager: blocking FIFO does not allow pushing -> underutilize CPU
% 
% The reason why Blocking Eager is usually slower than No Eager is due to \ttt{agg}regators that read multiple input streams at the same time---such \ttt{sort -m}.
% 
% Sort-sort possibly shows the full spectrum:
%   (iii) \sys's \ttt{eager} offers benefits despite the lack of \ttt{split} because it fully parallelizes the first \ttt{sort} while maintaining task parallelism
%   (iv) \sys gets full benefits because \ttt{split}ting allows parallelizing the second \ttt{sort} too.

      % (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      % Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
